---
title: "Create your own Model"
author: "Demo"
date: "August 22, 2018"
output: html_document
---

### Creating your Own Model
This Tutorial will demonstrate how to create a basic model with ForecastFramework. Models are created with R Object-Oriented programming, or the [R6 package](https://cran.r-project.org/web/packages/R6/vignettes/Introduction.html). It is essential to have a basic understanding of the structure of R6 Objects, before creating a ForecastFramework package. In this tutorial, the SARIMATD model will be created step-by-step.

### SARIMATD
SARIMATD is a model created by [Evan Ray](http://www.mtholyoke.edu/~eray/). SARIMATD uses set of wrapper functions around the forecast package to simplify estimation of and prediction from SARIMA models.

Basic Steps of the Model:

1. Incidence data are transformed to approximate normality
2. Seasonal differencing is performed
3. Transformed data are passed to `forecast::auto.arima` for model selection and estimation
simulate trajectories of future incidence

### Building the SARIMATD Model
#### 1. Start with an Empty Model
ForecastFramework models all have the same basic structure:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
model <- R6Class(
  inherit = ContestModel, # or AggregateModel
  private = list(
    #' private section is for aspects we do not want the user to access directly
  ),
  public = list(
    #' @method fit_ Get the model ready to forecast
    #' @param data The data to fit the model to.
    fit = function(data){
      ...
    },
    #' @method forecast Predict some number of time steps into the future.
    #' @param newdata The data to forecast from
    #' @param steps The number of timesteps into the future to forecast
    forecast = function(newdata,steps){
      ...
    },
    #' @method initialize Create a new instance of this class. 
    initialize = function(){
      ...
    }
  )
)
```

#### 2. Define your `fit()` and `forecast()` functions
The easiest way to incorporate a model into ForecastFramework is to pre-define your fit and forecast functions in a separate package, or use a separate package like the `forecast` package. The [SARMATD package](https://github.com/reichlab/sarimaTD) is hosted on GitHub and available for download with CRAN.

The SARIMATD fit function is as follows:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
#' @param y a univariate time series or numeric vector.
#' @param ts_frequency frequency of time series.  Must be provided if y is not
#'   of class "ts".  See the help for stats::ts for more.
#' @param transformation character specifying transformation type:
#'   "box-cox", "log", "forecast-box-cox", or "none".  See details for more.
#' @param seasonal_difference boolean; take a seasonal difference before passing
#'   to auto.arima?
#' @param auto.arima_d order of first differencing argument to auto.arima.
#' @param auto.arima_D order of seasonal differencing argument to auto.arima.
fit_sarima(
          y,
          ts_frequency,
          transformation = "box-cox",
          seasonal_difference = TRUE,
          d = NA,
          D = NA)
```

The SARIMATD forecast function is as follows (note that SARIMATD is a simulated forecast):
```{r, warning=FALSE, message=FALSE, eval=FALSE}
#' @param object a sarima fit of class "sarima_td", as returned by fit_sarima
#' @param nsim number of sample trajectories to simulate
#' @param seed either `NULL` or an integer that will be used in a call to
#'   `set.seed` before simulating the response vectors.  If set, the value is
#'   saved as the "seed" attribute of the returned value.  The default, `Null`,
#'   will not change the random generator state, and return `.Random.seed`
#'   as the "seed" attribute
#' @param newdata new data to simulate forward from
#' @param h number of time steps forwards to simulate
#'
#' @return an nsim by h matrix with simulated values
simulate.sarimaTD(
                  object,
                  nsim = 1,
                  seed = NULL,
                  newdata,
                  h = 1)
``` 

#### 3. Update the R6 Private Section
This section of the Model:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
 private = list(
    #' @method private section is for aspects we do not want the user to access directly
  ),
```
Will become:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
  private = list(
    .data = NULL,        ## every model should have this
    .models = list(),    ## specific to models that are fit separately for each location
    .nsim = 1000,        ## models that are simulating forecasts need this
    .lambda = list(),    ## specific to ARIMA models
    .period = integer(0) ## specific to SARIMA models
  ),
```
This section is comprised of inputs in your model that are not able to be accessed by the user directly, but will be updatable in the later initialize section. The arguments in the `fit()` and `forecast()` functions are fixed without incorporating private fields. For example, the `nsim` (number of simulations) may change, which makes it necessary to add these updatable variables in the private section. However, notice that all private variables include a `.` in the private section. This is the common way of distinguising a private variable in `R6`. 

#### 4. Update the fit Method
Upload IncidenceMatrix to Fit function:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
 fit = function(data) {
      ## All fit functions should have this, it helps with debugging
      if("fit" %in% private$.debug){browser()}
      
      ## stores data for easy access and checks to make sure it's the right class
      private$.data <- IncidenceMatrix$new(data)
      
 }
```

Cycle though each province and use the `fit_sarima()` function.
```{r, warning=FALSE, message=FALSE, eval=FALSE}
 fit = function(data) {
      ## All fit functions should have this, it helps with debugging
      if("fit" %in% private$.debug){browser()}
    
      ## stores data for easy access and checks to make sure it's the right class
      private$.data <- IncidenceMatrix$new(data)
      
      ## for each location/row
      for (row_idx in 1:private$.data$nrow) {
        ### need to create a y vector with incidence at time t
        y <- private$.data$subset(rows = row_idx, mutate = FALSE)
        
        ## private$.models[[row_idx]] <- something
        y_ts <- ts(as.vector(y$mat), frequency = private$.period)
        private$.lambda[[row_idx]] <- BoxCox.lambda(y_ts)
        private$.models[[row_idx]] <- fit_sarima(y = y_ts,
                                                 transformation = "box-cox",
                                                 seasonal_difference = TRUE)
      }
 }
```

#### 4. Update the forecast Method
Using the pre-defined equation `simulate.sarimaTD()` for the forecast method. `nmodels`, `sim_forecasts` and `dimnames(simforecasts)` are all variables needed to define a simulated forecast in ForecastFramework.
```{r, warning=FALSE, message=FALSE, eval=FALSE}
    forecast = function(newdata = private$.data, steps) {
      ## include for debugging
      if("forecast" %in% private$.debug){browser()} 
      
      ## number of models (provinces) to forecast
      nmodels <- length(private$.models)
      
      ## define an array to store the simulated forecasts
      sim_forecasts <- array(dim = c(nmodels, steps, private$.nsim))
      dimnames(sim_forecasts) <- list(newdata$rnames, 1:steps, NULL)
      
    }
```

Now, the Provinces are iterated through and simulated with the `simulate.sarimaTD()`function (called as just `simulate()`). Then, there are a couple of matrix transformations to input the simulated matrix into the `SimulatedIncidenceMatrix()` Object.
```{r, warning=FALSE, message=FALSE, eval=FALSE}
    forecast = function(newdata = private$.data, steps) {
      ## include for debugging
      if("forecast" %in% private$.debug){browser()} 
      
      ## number of models (provinces) to forecast
      nmodels <- length(private$.models)
      
      ## define an array to store the simulated forecasts
      sim_forecasts <- array(dim = c(nmodels, steps, private$.nsim))
      dimnames(sim_forecasts) <- list(newdata$rnames, 1:steps, NULL)
      
      ## iterate through each province and forecast with simulate.satimaTD
      for(model_idx in 1:length(private$.models)) {
          tmp_arima <-  simulate(object = private$.models[[model_idx]],
                                 nsim = private$.nsim,
                                 seed = 1,
                                 newdata = as.vector(newdata$mat[model_idx,]),
                                 h = steps
                                )
          ## transpose simulate() output to be consistent with ForecastFramework
          tmp_arima <- t(tmp_arima)
          sim_forecasts[model_idx, , ] <- tmp_arima
      }
      private$output <- SimulatedIncidenceMatrix$new(sim_forecasts)
      return(IncidenceForecast$new(private$output, forecastTimes = rep(TRUE, steps)))
    }
```

#### 5. Update the initialize Method
The initialize method is to initialize inputs. In this case, one input that needs initialization is the period (number of biweeks).
```{r, warning=FALSE, message=FALSE, eval=FALSE}
initialize = function(period = 26, nsim=1000) { 
      ## this code is run during SARIMAModel$new()
      ## need to store these arguments within the model object
      private$.nsim <- nsim
      private$.period <- period
    },
```

#### 6. Update the active binding
We will ignore this section for now. It is a more advanced level and stay tuned for another vignette that expains active binding.

#### 7. Putting everything Together
All of the previous sections and methods are put together as the R6 Object:

```{r, warning=FALSE, message=FALSE, eval=TRUE}
library(ForecastFramework)
library(R6)
library(forecast)
library(sarimaTD)

# SARIMATDModel Description:
# Implementation of the sarimaTD model created by Evan Ray
# More information about sarimaTD in https://github.com/reichlab/sarimaTD 
# Integrated into ForecastFramework by: Katie House, 6/22/2018

SARIMATD1Model <- R6Class(
  inherit = ContestModel,
  private = list(
    .data = NULL,        ## every model should have this
    .models = list(),    ## specific to models that are fit separately for each location
    .nsim = 1000,        ## models that are simulating forecasts need this
    .period = integer(0) ## specific to SARIMA models
  ),
  public = list(
    ## data will be MatrixData
    fit = function(data) {
      if("fit" %in% private$.debug){browser()}
      ## stores data for easy access and checks to make sure it's the right class
      private$.data <- IncidenceMatrix$new(data)
      
      ## for each location/row
      for (row_idx in 1:private$.data$nrow) {
        ### need to create a y vector with incidence at time t
        y <- private$.data$subset(rows = row_idx, mutate = FALSE)
        
        ## convert y vector to time series data type
        y_ts <- ts(as.vector(y$mat), frequency = private$.period)
        
        ## fit sarimaTD with 'fit_sarima()' from sarimaTD package
        ## fit_sarima() performs box-cox transformation and seasonal differencing
        private$.models[[row_idx]] <- fit_sarima(y = y_ts,
                                                 transformation = "box-cox",
                                                 seasonal_difference = TRUE)
      }
    },
    forecast = function(newdata = private$.data, steps) {
      ## include for debugging
      if("forecast" %in% private$.debug){browser()} 
      
      ## number of models (provinces) to forecast
      nmodels <- length(private$.models)
      
      ## define an array to store the simulated forecasts
      sim_forecasts <- array(dim = c(nmodels, steps, private$.nsim))
      dimnames(sim_forecasts) <- list(newdata$rnames, 1:steps, NULL)
      
      ## iterate through each province and forecast with simulate.satimaTD
      for(model_idx in 1:length(private$.models)) {
          tmp_arima <-  simulate(object = private$.models[[model_idx]],
                                 nsim = private$.nsim,
                                 seed = 1,
                                 newdata = as.vector(newdata$mat[model_idx,]),
                                 h = steps
                                )
          ## transpose simulate() output to be consistent with ForecastFramework
          tmp_arima <- t(tmp_arima)
          sim_forecasts[model_idx, , ] <- tmp_arima
      }
      private$output <- SimulatedIncidenceMatrix$new(sim_forecasts)
      return(IncidenceForecast$new(private$output, forecastTimes = rep(TRUE, steps)))
    },
    initialize = function(period = 26, nsim=1000) { 
      ## this code is run during SARIMAModel$new()
      ## need to store these arguments within the model object
      private$.nsim <- nsim
      private$.period <- period
    }
  )
)
```

#### 8. Testing the New Model
Lets now test the model on the same biweek counts as the SARIMA demo. The output is a forecasted case count for 26 biweeks.
```{r, warning=FALSE, message=FALSE}
# training data for province 10, years 2006 - 2012 
library(ForecastFramework)
library(R6)
library(forecast)
library(dplyr)
library(ggplot2)
library(cdcfluview)

# Source R6 Files
library(RCurl)

# Function of Source Github Models
source_github <- function(u) {
  script <- getURL(u, ssl.verifypeer = FALSE)
  eval(parse(text = script),envir=.GlobalEnv)
}  
# Source R6 Files
source_github('https://raw.githubusercontent.com/reichlab/forecast-framework-demos/master/models/ContestModel.R')
source_github('https://raw.githubusercontent.com/reichlab/forecast-framework-demos/master/models/SARIMATD1Model.R')

# Preprocess data to Inc Matrix
data2 <- ilinet(region = "National")
data2 <- data2 %>% filter( week != 53)
data2 <- data2 %>% 
  filter( year > 2009 & 
          year < 2018 & 
          !(year == 2017 & week > 18))
data2 <- data2 %>%
  select('region','year','week','weighted_ili', 'week_start')

preprocess_inc <- function(dat){
  dat$time.in.year = dat$week
  dat$t = dat$year + dat$week/52
  inc = ObservationList$new(dat)
  inc$formArray('region','t',val='weighted_ili',
                dimData = list(NULL,list('week','year','time.in.year','t')),
                metaData = list(t.step = 1/52,max.year.time = 52))
  return(inc)
}
training_data <- data2 %>% 
  filter( ! ((year == 2016 & week >= 19)| (year == 2017 ) ))
training_inc <- preprocess_inc(training_data)

# Create new SARIMATD model
nsim <- 10 # Number of SARIMA simulations 
sarimaTD_model <- SARIMATD1Model$new(period = 52, nsim = nsim)

# Fit SARIMATD model
sarimaTD_model$fit(training_inc)

# Forecast SARIMATD Model
steps <- 52 # forecast ahead 26 biweeks
forecast_X <- sarimaTD_model$forecast(steps = steps)

print(forecast_X$data$mat)
```