---
title: "Create your own Model"
author: "Demo"
date: "August 22, 2018"
output: html_document
---

### Creating your Own Model
This Tutorial will demonstrate how to create a basic model with ForecastFramework. Models are created with R Object-Oriented programming, or the [R6 package](https://cran.r-project.org/web/packages/R6/vignettes/Introduction.html). It is essential to have a basic understanding of the structure of R6 Objects, before creating a ForecastFramework package. In this tutorial, the SARIMATD model will be created step-by-step.

### SARIMATD
SARIMATD is a model created by Evan Ray of the Reich Lab. SARIMATD uses set of wrapper functions around the forecast package to simplify estimation of and prediction from SARIMA models.

Basic Steps of the Model:

1. Incidence data are transformed to approximate normality
2. Seasonal differencing is performed
3. Transformed data are passed to `forecast::auto.arima` for model selection and estimation
simulate trajectories of future incidence

### Building the SARIMATD Model
#### 1. Start with an Empty Model
ForecastFramework models all have the same basic structure:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
model <- R6Class(
  inherit = ContestModel, # or AggregateModel
  private = list(
    #' @method private section is for aspects we do not want the user to access directly
  ),
  public = list(
    #' @method fit_ Get the model ready to predict
    #' @param data The data to fit the model to.
    fit = function(data){
      ...
    },
    #' @method forecast Predict some number of time steps into the future.
    #' @param newdata The data to forecast from
    #' @param steps The number of timesteps into the future to predict.
    forecast = function(newdata,steps){
      ...
    },
    #' @method initialize Create a new instance of this class. 
    initialize = function(){
      ...
    },
    #' @method predict predict using the model.
    #' @param newdata Predict using the model.
    predict = function(newdata){
      ...
    },
  )
)
```

#### 2. Define your `fit()` and `forecast()` functions
The easiest way to incorporate a model into ForecastFramework is to pre-define your fit and forecast functions in a separate package, or use a separate package like the `forecast` package. The [SARMATD package](https://github.com/reichlab/sarimaTD) is hosted on Github and available for download with CRAN.

The SARIMATD fit function is as follows:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
#' @param y a univariate time series or numeric vector.
#' @param ts_frequency frequency of time series.  Must be provided if y is not
#'   of class "ts".  See the help for stats::ts for more.
#' @param transformation character specifying transformation type:
#'   "box-cox", "log", "forecast-box-cox", or "none".  See details for more.
#' @param seasonal_difference boolean; take a seasonal difference before passing
#'   to auto.arima?
#' @param auto.arima_d order of first differencing argument to auto.arima.
#' @param auto.arima_D order of seasonal differencing argument to auto.arima.
fit_sarima(
          y,
          ts_frequency,
          transformation = "box-cox",
          seasonal_difference = TRUE,
          d = NA,
          D = NA)
```

The SARIMATD forecast function is as follows (note that SARIMATD is a simulated forecast):
```{r, warning=FALSE, message=FALSE, eval=FALSE}
#' @param object a sarima fit of class "sarima_td", as returned by fit_sarima
#' @param nsim number of sample trajectories to simulate
#' @param seed either `NULL` or an integer that will be used in a call to
#'   `set.seed` before simulating the response vectors.  If set, the value is
#'   saved as the "seed" attribute of the returned value.  The default, `Null`,
#'   will not change the random generator state, and return `.Random.seed`
#'   as the "seed" attribute
#' @param newdata new data to simulate forward from
#' @param h number of time steps forwards to simulate
#'
#' @return an nsim by h matrix with simulated values
simulate.sarimaTD(
                  object,
                  nsim = 1,
                  seed = NULL,
                  newdata,
                  h = 1)
``` 

#### 3. Update the R6 Private Section
This section of the Model:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
 private = list(
    #' @method private section is for aspects we do not want the user to access directly
  ),
```
Will become:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
  private = list(
    .data = NULL,        ## every model should have this
    .models = list(),    ## specific to models that are fit separately for each location
    .nsim = 1000,         ## models that are simulating forecasts need this
    .lambda = list(),    ## specific to ARIMA models
    .period = integer(0) ## specific to SARIMA models
  ),
```

#### 4. Update the fit Method
Upload IncidenceMatrix to Fit function:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
 fit = function(data) {
      ## All fit functions should have this, it helps with debugging
      if("fit" %in% private$.debug){browser()}
      
      ## stores data for easy access and checks to make sure it's the right class
      private$.data <- IncidenceMatrix$new(data)
      
 }
```

Cycle though each province and use the `fit_sarima()` function.
```{r, warning=FALSE, message=FALSE, eval=FALSE}
 fit = function(data) {
      ## All fit functions should have this, it helps with debugging
      if("fit" %in% private$.debug){browser()}
    
      ## stores data for easy access and checks to make sure it's the right class
      private$.data <- IncidenceMatrix$new(data)
      
      ## for each location/row
      for (row_idx in 1:private$.data$nrow) {
        ### need to create a y vector with incidence at time t
        y <- private$.data$subset(rows = row_idx, mutate = FALSE)
        
        ## private$.models[[row_idx]] <- something
        y_ts <- ts(as.vector(y$mat), frequency = private$.period)
        private$.lambda[[row_idx]] <- BoxCox.lambda(y_ts)
        private$.models[[row_idx]] <- fit_sarima(y = y_ts,
                                                 transformation = "box-cox",
                                                 seasonal_difference = TRUE)
      }
 }
```

#### 4. Update the forecast Method
Using the pre-defined equation `simulate.sarimaTD()` for the forecast method. `nmodels`, `sim_forecasts` and `dimnames(simforecasts)` are all variables needed to define a simulated forecast in ForecastFramework.
```{r, warning=FALSE, message=FALSE, eval=FALSE}
    forecast = function(newdata = private$.data, steps) {
      ## include for debugging
      if("forecast" %in% private$.debug){browser()} 
      
      ## number of models (provinces) to forecast
      nmodels <- length(private$.models)
      
      ## define an array to store the simulated forecasts
      sim_forecasts <- array(dim = c(nmodels, steps, private$.nsim))
      dimnames(sim_forecasts) <- list(newdata$rnames, 1:steps, NULL)
      
    }
```

Now, the Provinces are iterated through and simulated with the `simulate.sarimaTD()`function (called as just `simulate()`). Then, there are a couple of matrix transformations to input the simulated matrix into the `SimulatedIncidenceMatrix()` Object.
```{r, warning=FALSE, message=FALSE, eval=FALSE}
    forecast = function(newdata = private$.data, steps) {
      ## include for debugging
      if("forecast" %in% private$.debug){browser()} 
      
      ## number of models (provinces) to forecast
      nmodels <- length(private$.models)
      
      ## define an array to store the simulated forecasts
      sim_forecasts <- array(dim = c(nmodels, steps, private$.nsim))
      dimnames(sim_forecasts) <- list(newdata$rnames, 1:steps, NULL)
      
      ## iterate through each province and forecast with simulate.satimaTD
      for(model_idx in 1:length(private$.models)) {
          tmp_arima <-  simulate(object = private$.models[[model_idx]],
                                 nsim = private$.nsim,
                                 seed = 1,
                                 newdata = as.vector(newdata$mat[model_idx,]),
                                 h = steps
                                )
          ## transpose simulate() output to be consistent with ForecastFramework
          tmp_arima <- t(tmp_arima)
          sim_forecasts[model_idx, , ] <- tmp_arima
      }
      private$output <- SimulatedIncidenceMatrix$new(sim_forecasts)
      return(IncidenceForecast$new(private$output, forecastTimes = rep(TRUE, steps)))
    }
```

#### 5. Update the initialize Method
The initialize method is to initialize inputs. In this case, one input that needs initialization is the period (number of biweeks).
```{r, warning=FALSE, message=FALSE, eval=FALSE}
initialize = function(period = 26, nsim=1000) { 
      ## this code is run during SARIMAModel$new()
      ## need to store these arguments within the model object
      private$.nsim <- nsim
      private$.period <- period
    },
```

#### 6. Update the active section
This section determines what users can change in the model:
```{r, warning=FALSE, message=FALSE, eval=FALSE}
active = list(
    ## This list determines how you can access pieces of your model object
    data = function(value) {
      ## use this form when you want this parameter to be un-modifiable
      if(!missing(value))
        stop("Writing directly to the data is not allowed.")
      return(private$.data)
    },
    models = function(value) {
      ## use this form when you want this parameter to be un-modifiable
      if(!missing(value))
        stop("Writing directly to the models is not allowed.")
      return(private$.models)
    },
    nsim = function(value) {
      ## use this form when you want to be able to change this parameter
      private$defaultActive(type="private", ".nsim", val=value)
    },
    lambda = function(value) {
      ## use this form when you want this parameter to be un-modifiable
      if(!missing(value))
        stop("Writing directly to lambda is not allowed.")
      return(private$.lambda)
    },
    period = function(value) {
      ## use this form when you want this parameter to be un-modifiable
      if(!missing(value))
        stop("Writing directly to the model period is not allowed.")
      return(private$.period)
    }
  )
```

#### 7. Putting everything Together
All of the previous sections and methods are put together as the R6 Object:

```{r, warning=FALSE, message=FALSE, eval=TRUE}
library(ForecastFramework)
library(R6)
library(forecast)
library(sarimaTD)

# SARIMATDModel Description:
# Implementation of the sarimaTD model created by Evan Ray
# More information about sarimaTD in https://github.com/reichlab/sarimaTD 
# Integrated into ForecastFramework by: Katie House, 6/22/2018

SARIMATD1Model <- R6Class(
  inherit = ContestModel,
  private = list(
    .data = NULL,        ## every model should have this
    .models = list(),    ## specific to models that are fit separately for each location
    .nsim = 1000,        ## models that are simulating forecasts need this
    .period = integer(0) ## specific to SARIMA models
  ),
  public = list(
    ## data will be MatrixData
    fit = function(data) {
      if("fit" %in% private$.debug){browser()}
      ## stores data for easy access and checks to make sure it's the right class
      private$.data <- IncidenceMatrix$new(data)
      
      ## for each location/row
      for (row_idx in 1:private$.data$nrow) {
        ### need to create a y vector with incidence at time t
        y <- private$.data$subset(rows = row_idx, mutate = FALSE)
        
        ## convert y vector to time series data type
        y_ts <- ts(as.vector(y$mat), frequency = private$.period)
        
        ## fit sarimaTD with 'fit_sarima()' from sarimaTD package
        ## fit_sarima() performs box-cox transformation and seasonal differencing
        private$.models[[row_idx]] <- fit_sarima(y = y_ts,
                                                 transformation = "box-cox",
                                                 seasonal_difference = TRUE)
      }
    },
    forecast = function(newdata = private$.data, steps) {
      ## include for debugging
      if("forecast" %in% private$.debug){browser()} 
      
      ## number of models (provinces) to forecast
      nmodels <- length(private$.models)
      
      ## define an array to store the simulated forecasts
      sim_forecasts <- array(dim = c(nmodels, steps, private$.nsim))
      dimnames(sim_forecasts) <- list(newdata$rnames, 1:steps, NULL)
      
      ## iterate through each province and forecast with simulate.satimaTD
      for(model_idx in 1:length(private$.models)) {
          tmp_arima <-  simulate(object = private$.models[[model_idx]],
                                 nsim = private$.nsim,
                                 seed = 1,
                                 newdata = as.vector(newdata$mat[model_idx,]),
                                 h = steps
                                )
          ## transpose simulate() output to be consistent with ForecastFramework
          tmp_arima <- t(tmp_arima)
          sim_forecasts[model_idx, , ] <- tmp_arima
      }
      private$output <- SimulatedIncidenceMatrix$new(sim_forecasts)
      return(IncidenceForecast$new(private$output, forecastTimes = rep(TRUE, steps)))
    },
    initialize = function(period = 26, nsim=1000) { 
      ## this code is run during SARIMAModel$new()
      ## need to store these arguments within the model object
      private$.nsim <- nsim
      private$.period <- period
    },
    predict = function(newdata) {
      stop("predict method has not been written.")
    }
  ),
  active = list(
    ## This list determines how you can access pieces of your model object
    data = function(value) {
      ## use this form when you want this parameter to be un-modifiable
      if(!missing(value))
        stop("Writing directly to the data is not allowed.")
      return(private$.data)
    },
    models = function(value) {
      ## use this form when you want this parameter to be un-modifiable
      if(!missing(value))
        stop("Writing directly to the models is not allowed.")
      return(private$.models)
    },
    nsim = function(value) {
      ## use this form when you want to be able to change this parameter
      private$defaultActive(type="private", ".nsim", val=value)
    },
    period = function(value) {
      ## use this form when you want this parameter to be un-modifiable
      if(!missing(value))
        stop("Writing directly to the model period is not allowed.")
      return(private$.period)
    }
  )
)
```

#### 8. Testing the New Model
Lets now test the model on the same biweek counts as the SARIMA demo. The output is a forecasted case count for 26 biweeks.
```{r, warning=FALSE, message=FALSE}
# training data for province 10, years 2006 - 2012 
library(ForecastFramework)
library(R6)
library(forecast)
library(dplyr)
library(ggplot2)
library(cdcfluview)

# Source R6 Files
library(RCurl)

# Function of Source Github Models
source_github <- function(u) {
  script <- getURL(u, ssl.verifypeer = FALSE)
  eval(parse(text = script),envir=.GlobalEnv)
}  
# Source R6 Files
source_github('https://raw.githubusercontent.com/reichlab/forecast-framework-demos/master/models/ContestModel.R')
source_github('https://raw.githubusercontent.com/reichlab/forecast-framework-demos/master/models/SARIMATD1Model.R')

# Preprocess data to Inc Matrix
data2 <- ilinet(region = "National")
data2 <- data2 %>% filter( week != 53)
data2 <- data2 %>% 
  filter( year > 2009 & 
          year < 2018 & 
          !(year == 2017 & week > 18))
data2 <- data2 %>%
  select('region','year','week','weighted_ili', 'week_start')

preprocess_inc <- function(dat){
  dat$time.in.year = dat$week
  dat$t = dat$year + dat$week/52
  inc = ObservationList$new(dat)
  inc$formArray('region','t',val='weighted_ili',
                dimData = list(NULL,list('week','year','time.in.year','t')),
                metaData = list(t.step = 1/52,max.year.time = 52))
  return(inc)
}
training_data <- data2 %>% 
  filter( ! ((year == 2016 & week >= 19)| (year == 2017 ) ))
training_inc <- preprocess_inc(training_data)

# Create new SARIMATD model
nsim <- 10 # Number of SARIMA simulations 
sarimaTD_model <- SARIMATD1Model$new(period = 52, nsim = nsim)

# Fit SARIMATD model
sarimaTD_model$fit(training_inc)

# Forecast SARIMATD Model
steps <- 52 # forecast ahead 26 biweeks
forecast_X <- sarimaTD_model$forecast(steps = steps)

print(forecast_X$data$mat)
```