---
title: "The Data"
author: "Demo"
date: "August 22, 2018"
output: html_document
---

### Demo Data: CDC U.S. Flu Season Data
This demonstration uses [influenza surveillance data](https://www.cdc.gov/flu/weekly/overview.htm) maintained by the U.S. Centers for Disease Control (CDC). Normally, the data include all 50 US states. For simplicity, this demo only models national data. These data are quickly retrieved with the [`cdcfluview` R package](https://github.com/hrbrmstr/cdcfluview).

One of the measures reported by the CDC for quantifying respiratory illness burden in a population is weighted influenza-like illness (wILI). This is a metric defined as the estimated percentage of all out-patient doctors office visits due to patients who present with "influenza-like illness", e.g. a fever, a cough and/or sore throat. This wILI percentage, at the national and regional levels, is calculated by taking weighted averages of state-level activty, with weights based on state population.

#### Raw Data
These data include many attributes, but we will focus on: `region`,`year`,`week`,`weighted_ili`,`week_start`. Note that `weighted_ili` is the wILI measure reported for that specific year and week.

```{r, warning=FALSE, message=FALSE}
library(cdcfluview)
library(dplyr)
dat <- ilinet(region = "National")
dat <- dat %>%
  select('region','year','week','weighted_ili', 'week_start')
print(head(dat,6))
```

#### Time Series
A time series of the raw data looks like the following:
```{r,fig.width=7, fig.height=4, warning=FALSE, message=FALSE}
library(ggplot2)
dat$date_sick <- as.Date(strptime(dat$week_start,"%m/%d/%Y")) # convert to date
plot <- ggplot() +
    geom_line(mapping = aes(x = week_start, y = weighted_ili),
              size=0.7,
              data = dat) +
    xlab("") + ylab("Weighted ILI") +
    coord_cartesian(ylim = c(0, 10)) +
    ggtitle("National Weighted Inluenza-Like Illness")
print(plot)
```

#### Scope of this demo
This demo will forecast an entire year's worth of wILI incidence starting from the beginning of the flu season in 2016.


<a href="#defining-inputs-incidence-matrix-1"><button type="button" class="btn btn-raised btn-secondary">Proceed to Defining the Inputs</button></a>