---
title: "SARIMA Demo"
author: "Katie House"
date: "June 20, 2018"
output: html_document
---

This demonstration investigate the usage of models with ForecastFramework. For more information on the SARIMA Model itself, check out: <https://github.com/reichlab/thai-dengue-district-challenge/blob/master/doc/ff-intro.Rmd>.

### Fitting the ForecastFramework SARIMA Model
#### 1. Import packages
First you must import all the required libraries. Note that ForecastFramework doesn't requre `dplyr` or `ggplot2`, but they will be used to make a figure at the end of the demo.

```{r, warning=FALSE, message=FALSE}
require(ForecastFramework)
require(R6)
require(forecast)
require(dplyr)
require(ggplot2)

# Source R6 Files
source('../models/ContestModel.R')
source('../models/SARIMAModel.R')
```

Now it is time to fit your SARIMA Model! ForecastFramework makes this easy. 

#### 2. Create new sarima_model Class
In this SARIMA model, simulation is used with `::forecast auto.arima()`. So we must define the number of simulations to use, `nsim`. We also need to define the seasonal periodicity of our data, or the `period` parameter. For Dengue forecasting, we know the periodicity is 26, because the data is biweekly. once we define these parameters, they can be passed into the `SARIMAModel()` class generator to create a new `sarima_model` class.

```{r, warning=FALSE, message=FALSE}
nsim <- 1000 # Number of SARIMA simulations 
sarima_model <- SARIMAModel$new(period = 26, nsim = nsim)
```

#### 3. Import Dengue Data
To import data from a local csv, use the `read.csv()` command. In this example, the `date_sick` field is imported as a wrong datatype, so it must be converted to a date value with a couple of transformations. Then, the data is stored as an `IncidenceMatrix()`. An incidence matrix represents spatial time-series data in a matrix with one row per location and one column per time. To do this, another series of transformations are made. IncidenceMatrix is the preferred input for ForecastFramework Models.
```{r, warning=FALSE, message=FALSE}
# training data for province 10, years 2006 - 2012 
dat <- read.csv('../data/province-biweek-counts-training.csv')
dat$date_sick <- as.Date(strptime(dat$date_sick,"%m/%d/%Y")) # convert to date values
inc <- IncidenceMatrix$new(1+reshape2::acast(dat,province~date_sick,value.var='cases'))
print(inc$mat[,0:10])
```

Let's look at a sample from the new IncidenceMatrix to view the IncidenceMatrixx Format:
```{r, warning=FALSE, message=FALSE}
print(inc$mat[,0:10])
```

#### 3. Fit the Model
Typically, all the provinces in Thailand would be fit separately and have different forecasts. For the scope of this demo, we are only fitting and forecasting Province 10, or Bangkok Metropolis. The ForecastFramework SARIMA Model method requires the user to specify which province to model, in this case, we are modeling the first (and only) province in the dataset.

```{r, warning=FALSE, message=FALSE}
# define how many provinces to model
# this demo only has data for one province: Province 10
prov_nums <- 1
nmodels <- length(prov_nums)
sarima_model$fit(inc$subset(rows = prov_nums, mutate = FALSE))
```

### Forecasting with the ForecastFramework SARIMA Model
#### 4. Forecasting SARIMA
Before forecasting, you must specify how many periods ahead you would like to forecast. This is defined by the `steps` parameter. Then, you forecast with the `sarima_model$forecast()` method. This model will forecast one year into the future (26 biweeks).

```{r, warning=FALSE, message=FALSE}
# define how many provinces to model
# this demo only has data for one province: Province 10
steps <- 26 # forecast ahead 26 biweeks
```

Then, use the built-in `$forecast()` function to create your forecasts. This creates a new R6 object with your forecasts and many other useful built-in functions. Be sure to check out the Evaluation section for usage of the Forecasting function.
```{r, warning=FALSE, message=FALSE}
forecast_X <- sarima_model$forecast(steps = steps)
```

### Plotting the Output
#### 5. Formatting the Forecast Output
To view the new forecasting data in matrix form, execute the following command:
```{r, warning=FALSE}
forecast_X$data$mat
```

Lets convert the forecast matrix to a dataframe in R for easy manipulation: 
```{r, warning=FALSE, message=FALSE}
# converting predictions to a dataframe to use dplyr
preds_df <- data.frame(as.table(t(forecast_X$data$mat)))
```

To include the `date_sick` in the  `forecast_X$data$mat` output, map the forecast to its predicted dates is to create a new data frame with the data from the missing year. Then, we will combine the output in `forecast_X$data` with the correct biweek dates.
```{r, warning=FALSE, message=FALSE}
# converting predictions to a dataframe to use dplyr
# import testing dataset which has 2013 data
data_X <- read.csv('../data/province-biweek-counts-testing.csv')
data_X$date_sick <- as.Date(strptime(data_X$date_sick,"%m/%d/%Y")) # convert to date values
preds_dates <- data_X %>%
          filter(year == 2013)
# add prediction dates to original forecast
preds_df[["date_sick"]] <-as.Date(preds_dates$date_sick, format = "%d-%m-%Y")
```

Now, we can add the forecast to the `preds_df`. Notice the `$median()` and `$quantile()` functions accompanied with `$mat` will produce matricies with the prediction medians and quantiles, respectively. 
```{r, warning=FALSE, message=FALSE}
preds_df <- preds_df %>%
   mutate( 
      pred_total_cases = as.vector(forecast_X$median()$mat),
      pred_95_lb = as.vector(forecast_X$quantile(0.025)$mat),
      pred_95_ub = as.vector(forecast_X$quantile(0.975)$mat),
      pred_80_lb = as.vector(forecast_X$quantile(0.05)$mat),
      pred_80_ub = as.vector(forecast_X$quantile(0.95)$mat),
      pred_50_lb = as.vector(forecast_X$quantile(0.25)$mat),
      pred_50_ub = as.vector(forecast_X$quantile(0.75)$mat)
   )
print(head(preds_df,5)) # Print the first 5 rows
```

#### 6. Plot the Forecast Output
Finally, you are able to plot your SARIMA Forecast! Lets use `ggplot2` to create a cool plot.
### Bangkok Metropolis SARIMA Forecast
```{r, warning=FALSE, message=FALSE,echo=FALSE,fig.width=10, fig.height=4}
ggplot() +
  geom_ribbon(
    mapping = aes(x = date_sick, ymin = pred_95_lb, ymax = pred_95_ub),
    fill = "cornflowerblue",
    alpha = 0.2,
    data = preds_df) +
  geom_ribbon(
    mapping = aes(x = date_sick, ymin = pred_80_lb, ymax = pred_80_ub),
    fill = "cornflowerblue",
    alpha = 0.2,
    data = preds_df) +
  geom_ribbon(
    mapping = aes(x = date_sick, ymin = pred_50_lb, ymax = pred_50_ub),
    alpha = 0.2,
    data = preds_df) +
  geom_line(
    mapping = aes(x = date_sick, y = pred_total_cases),
    color = "royalblue",
    size = 2,
    data = preds_df) +
  geom_line(mapping = aes(x = date_sick, y = cases),
            size=0.7,
            data = data_X) +
  xlab("") + ylab("Number of Cases") +
  coord_cartesian(ylim = c(0, 1000))
```