---
title: "evaluation"
author: "Demo"
date: "August 21, 2018"
output: html_document
---

### Evaluating Multiple Models
This tutorial will demonstrate how to easily compare multiple models on ForecastFramework. One of the best advantages of ForecastFramework is the ability to evaluate many models at once with the same functions.

#### Preprocessing Data
Import libraries and source models
```{r, warning=FALSE, message=FALSE}
library(ForecastFramework)
library(R6)
library(forecast)
library(dplyr)
library(ggplot2)
library(cdcfluview)
library(gridExtra)
library(data.table)
library(knitr)
library(kableExtra)
library(RCurl)

# Function of Source Github Models
source_github <- function(u) {
  # read script lines from website and evaluate
  script <- getURL(u, ssl.verifypeer = FALSE)
  eval(parse(text = script),envir=.GlobalEnv)
}  
# Source R6 Files
source_github('https://raw.githubusercontent.com/reichlab/forecast-framework-demos/master/models/ContestModel.R')
source_github('https://raw.githubusercontent.com/reichlab/forecast-framework-demos/master/models/AggregateForecastModel.R')
source_github('https://raw.githubusercontent.com/reichlab/forecast-framework-demos/master/models/SARIMAModel.R')
source_github('https://raw.githubusercontent.com/reichlab/forecast-framework-demos/master/models/GamModel.R')
source_github('https://raw.githubusercontent.com/reichlab/forecast-framework-demos/master/models/SARIMATD1Model.R')
```

Upload data from cdcfluview as dataframes. In this demo, we are only concerned with years between 2010 and 2017. We split training/testing data by excluding the 2016-17 Flu season in the training set. 
```{r, warning=FALSE, message=FALSE}
data <- ilinet(region = "National")
data <- data %>% filter( week != 53)
data <- data %>% 
  filter( year > 2009 & year < 2018 & 
          !(year == 2017 & week > 18))
data <- data %>%
  select('region','year','week','weighted_ili', 'week_start')

# Seperate into Training and Testing Data
training_data <- data %>% 
  filter( ! ((year == 2016 & week >= 19)| (year == 2017 ) ))
testing_data <- data %>% 
  filter((year == 2016 & week >= 19) | (year == 2017 & week <= 18) )
```

A **sample of the data** we are forecasting.
```{r, warning=FALSE, message=FALSE}
head(testing_data,5)
```

Then, preprocess both the training and testing dataframes to become new incidence matrices. This script uses the `ObservationList$new()` ForecastFramework function to easily convert the dataframe to an `IncidenceMatrix`. The `$formArray()` is the second step to complete the `IncidenceMatrix`.

Note: The time step in this example is 52 weeks; however, there are datasets with other timesteps (e.g. 26 biweeks) that would impact the calculations of `t.step`, `dat$t`, and `max.year.time` values.
```{r, warning=FALSE, message=FALSE}
# Create a Preprocess Function
# Convert Training and Testing data to incidence matrices
# time.in.year is a requirement for the GAM model
time_step <- 52
preprocess_inc <- function(dat){
  dat$time.in.year = dat$week
  dat$t = dat$year + dat$week/time_step
  inc = ObservationList$new(dat)
  inc$formArray('region','t',val='weighted_ili',
                dimData = list(NULL,list('week','year','time.in.year','t')),
                metaData = list(t.step = 1/time_step, max.year.time = time_step))
  return(inc)
}

# Preprocess Data and convert to Inc Mat
training_inc <- preprocess_inc(training_data)
testing_inc <- preprocess_inc(testing_data)
```

#### Defining Models
This demo will be evaluating the Generalized Additive Model (GAM), SARIMA, and sarimaTD models given different evaluation metrics. The GAM model is based on the model used in [published real-time infectious disease forecasting research](https://journals.plos.org/plosntds/article?id=10.1371/journal.pntd.0004761).

```{r, warning=FALSE, message=FALSE}
# Defining the Models
nsim <- 1000 # Number of simulations 
region_num <- 1 # Only modeling the "National" region

sarimaTD_model <- SARIMATD1Model$new(period = time_step, nsim = nsim)
sarima_model <- SARIMAModel$new(period = time_step, nsim = nsim)
gam_model <- GamModel$new(numProvinces = region_num, nSims = nsim)
```

#### Fitting Models
Each model is fit with the training data.

```{r, warning=FALSE, message=FALSE}
# Fit the Models 
sarimaTD_model$fit(training_inc)
sarima_model$fit(training_inc)
gam_model$fit(training_inc)
```

#### Forecasting Models
We are forecasting from the beginning of flu season 2016-17, 52 weeks into the future.
```{r, warning=FALSE, message=FALSE}
# Forecasting the Models
steps <- 52 # forecast ahead 26 biweeks
forecast_sarimaTD <- sarimaTD_model$forecast(step = steps)
forecast_sarima <- sarima_model$forecast(step = steps)
forecast_gam <- gam_model$forecast(step = steps)
```

#### Evaluating the Models
Now for the fun part! How did the models perform? This question is easy to answer with ForecastFramework Models.

**Important Note:** This is not a formal model evaluation or comparison. The goal of this exercise is to demonstrate different ForecastFramework techniques. Although the sarimaTD model appears to perform better than the other models in this example, a more in-depth analysis must be made before any conclusions can be made about model performance.

**Side-by-Side Forecasting Timeseries**
This function creates a time series visualization with the predicted median and quantiles displayed by a specific color. Notice `forecast$quantile(0.025)$mat` gives a matrix with a 2.5% quantile. Additionally, `forecast$mean()$mat` gives the mean forecast for that specific time period. Additionally, `na.rm=TRUE`, removes any NA values in the simulation to calculate the mean.

```{r, warning=FALSE, message=FALSE,fig.width=11, fig.height=5,}
# Function to create time series with Visualization
visualize_model <- function(forecast, rib_color, main_color){
  data_X_3years <- data %>% filter(year > 2013)
  preds_df <- data.frame(as.table(t(forecast$data$mat)))
  preds_df[["week_start"]] <-as.Date(testing_data$week_start, format = "%Y-%d-%m")
  preds_df <- preds_df %>%
    mutate( 
      pred_total_cases = as.vector(forecast$mean(na.rm=TRUE)$mat),
      pred_95_lb = as.vector(forecast$quantile(0.025,na.rm=TRUE)$mat),
      pred_95_ub = as.vector(forecast$quantile(0.975,na.rm=TRUE)$mat),
      pred_80_lb = as.vector(forecast$quantile(0.05,na.rm=TRUE)$mat),
      pred_80_ub = as.vector(forecast$quantile(0.95,na.rm=TRUE)$mat),
      pred_50_lb = as.vector(forecast$quantile(0.25,na.rm=TRUE)$mat),
      pred_50_ub = as.vector(forecast$quantile(0.75,na.rm=TRUE)$mat)
    )
  plot <- ggplot() +
    geom_ribbon(
      mapping = aes(x = week_start, ymin = pred_95_lb, ymax = pred_95_ub),
      fill = rib_color,
      alpha = 0.2,
      data = preds_df) +
    geom_ribbon(
      mapping = aes(x = week_start, ymin = pred_80_lb, ymax = pred_80_ub),
      fill = rib_color,
      alpha = 0.2,
      data = preds_df) +
    geom_ribbon(
      mapping = aes(x = week_start, ymin = pred_50_lb, ymax = pred_50_ub),
      fill = rib_color,
      alpha = 0.2,
      data = preds_df) +
    geom_line(
      mapping = aes(x = week_start, y = pred_total_cases),
      color = main_color,
      size = 1,
      data = preds_df) + 
    geom_line(mapping = aes(x = week_start, y = weighted_ili),
              size=0.7,
              data = data_X_3years) +
    xlab("") + ylab("Weighted ILI") + coord_cartesian(ylim = c(0, 13))
  return(plot)
}

# Create each plot
plot1 <- visualize_model(forecast_sarimaTD, "palegreen3", "forestgreen") + ggtitle("sarimaTD")
plot2 <- visualize_model(forecast_sarima, "coral1", "red") + ggtitle("SARIMA")
plot3 <- visualize_model(forecast_gam, "cornflowerblue" , "blue")+ ggtitle("GAM")

# Display side-by-side plots
grid.arrange(plot1, plot2, plot3, ncol=3)
```
Note the sparsity in the SARIMA forecast, due to some forecasted trajectories yielding undefined forecasted means (e.g. if one of the trajectories contains a non-numeric or NA value). This is one benefit of sarimaTD - the ability to provide stable long-term forecasts with the aid of transformations and seasonal differencing. It avoids some of the "brittlenes" that we have observed when just using SARIMA fitting out of the box. 

**Common Evaluation Metrics**
Calculating evaluation metrics is easy with ForecastFramework objects.
```{r, warning=FALSE, message=FALSE}
# Mean Absolute Error
MAE <- function(forecast_mean, test_inc){
  err <- test_inc$mat-forecast_mean
  abs_err <- abs(err)
  return(mean(abs_err))
}
# Mean Squared Error
MSE <- function(forecast_mean, test_inc){
  err <- test_inc$mat-forecast_mean
  err_sq <- err^2
  return(mean(err_sq))
}
# Mean Absolute Percentage Error
MAPE <- function(forecast_mean, test_inc){
  err <- test_inc$mat-forecast_mean
  abs_err <- abs(err)
  abs_err_percent <- abs_err/test_inc$mat
  return(mean(abs_err_percent)*100)
}
# Root Mean Squared Error
RMSE <- function(forecast_mean, test_inc){
  err <- test_inc$mat-forecast_mean
  err_sq <- err^2
  mse <-mean(err_sq)
  return(sqrt(mse))
}
```

Each Model can be evaluated given the various functions:
```{r, warning=FALSE, message=FALSE}
# Create Arrays with each Metric
evaluate <- function(forecast){
  # Replace any NaN values with 0
  forecast_mat <- forecast$mean(na.rm=TRUE)$mat
  forecast_mat[is.na(forecast_mat)] <- 0
  
  # Evaluate Performance
  MAE <- MAE(forecast_mat, testing_inc)
  MSE <- MSE(forecast_mat, testing_inc)
  MAPE <- MAPE(forecast_mat, testing_inc)
  RMSE <- RMSE(forecast_mat, testing_inc)
  return(c(MAE,MSE,MAPE,RMSE))
}

gam_eval <- evaluate(forecast_gam)
sarimaTD_eval <- evaluate(forecast_sarimaTD)
sarima_eval <- evaluate(forecast_sarima)
```

Then, a final Table is made with all of the metrics in a grid format:
```{r echo = FALSE, results = 'asis'}
df <- data.frame("MAE" = c(sarimaTD_eval[1],sarima_eval[1],gam_eval[1]),
                 "MSE" = c(sarimaTD_eval[2],sarima_eval[2],gam_eval[2]),
                 "MAPE" = c(sarimaTD_eval[3],sarima_eval[3],gam_eval[3]),
                 "RMSE" = c(sarimaTD_eval[4],sarima_eval[4],gam_eval[4]),
                 stringsAsFactors = FALSE)
row.names(df) <- c('sarimaTD','SARIMA','GAM')
results <- t(df)
kable(results) %>%
  kable_styling( bootstrap_options = "striped", 
                 full_width = F, position = "center")
```


<a href="#creating-your-own-model"><button type="button" class="btn btn-raised btn-secondary">Proceed to Creating your own Model</button></a>